---
title: "Chapter 32: Exponential Random Variables"
author: "Meike Niederhausen and Nicky Wakim"
date: "11/27/2023"
title-slide-attributes:
    data-background-color: "#006a4e"
categories: ["Week 10"]
format: 
  revealjs:
    theme: [default, simple_NW.scss]
    toc: true
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: Chapter 32 Slides
---

## Learning Objectives

1.  Identify the variable and the parameters in a story, and state in English what the variable and its parameters mean.

2.  Use the formulas for the density, CDF, expected value, and variance to answer questions and find probabilities.

```{css}
code.sourceCode {
  font-size: 1.5em;
  /* or try font-size: xx-large; */
}
```

## Properties of exponential RVs

-   **Scenario:** Modeling the time until the next (first) event

-   Continuous analog to the geometric distribution!

-   Shorthand: $X \sim \text{Exp}(\lambda)$

$$
f_X(x) = \lambda e^{-\lambda x}\text{ for } x>0, \lambda>0
$$

$$
F_X(x) = \left\{
        \begin{array}{ll}
            0 & \quad x<0 \quad \\
            1 - e^{-\lambda x} & \quad x\geq0 \\
        \end{array}
    \right.
$$

$$\text{E}(X) = \dfrac{1}{\lambda}$$ $$\text{Var}(X) = \dfrac{1}{\lambda^2}$$

## Memoryless Property

   

If $b>0$,

$$P(X > a +b | X> a) = P(X > b)$$

     

-   **This can be interpreted as:**

    -   If you have waited $a$ seconds (or any other measure of time) without a success

    -   Then the probability that you have to wait $b$ more seconds is the same as as the probability of waiting $b$ seconds initially.

## Identifying exponential RV from word problems

-   Look for time between events/successes

-   Look for a rate of the events over time period

-   How does it differ from the geometric distribution?

    -   Geometric is *number of trials* until first success

    -   Exponential is *time* until first success

-   Relation to the Poisson distribution?

    -   When the time between arrivals is exponential, the number of arrivals in a fixed time interval is Poisson with the mean $\lambda$

## Helpful R code

Let's say we're sitting at the bus stop, measuring the time until our bus arrives. We know the bus comes every 10 minutes on average.

-   If we want to know the probability that the bus arrives in the next 5 minutes:

    ```{r}
    #| echo: true
    pexp(q = 5, rate = 1/10)
    ```

-   If we want to know the time, say $t$, where the probability of the bus arriving at $t$ or earlier is 0.35:

    ```{r}
    #| echo: true
    qexp(p = 0.35, rate = 1/10)
    ```

-   If we want to know the probability that the bus arrives between 3 and 5 minutes:

    ```{r}
    #| echo: true
    pexp(q = 5, rate = 1/10) - pexp(q = 3, rate = 1/10)
    ```

-   If we want to sample 20 bus arrival times from the distribution:

    ```{r}
    #| echo: true
    rexp(n = 20, rate = 1/10)
    ```

## Transformation of independent exponential RVs

::: columns
::: {.column width="30%"}
::: example
::: ex-title
Example 1
:::

::: ex-cont
Let $X_i \sim \textrm{Exp}(\lambda_i)$ be independent RVs, for $i=1 \ldots n$. Find the pdf for the first of the arrival times.
:::
:::
:::
:::
