---
title: "Chapter 33: Gamma Random Variables"
author: "Meike Niederhausen and Nicky Wakim"
date: "11/29/2023"
title-slide-attributes:
    data-background-color: "#006a4e"
categories: ["Week 10"]
format: 
  revealjs:
    theme: [default, simple_NW.scss]
    toc: true
    chalkboard: true
    slide-number: true
    show-slide-number: all
    width: 1955
    height: 1100
    footer: Chapter 33 Slides
---

## Learning Objectives

1.  Identify the variable and the parameters in a story, and state in English what the variable and its parameters mean.

```{css}
code.sourceCode {
  font-size: 1.5em;
  /* or try font-size: xx-large; */
}
```

## Properties of gamma RVs

-   **Scenario:** Modeling the time until the $r^{th}$ event.
-   Continuous analog to the Negative Binomial distribution
-   Shorthand: $X \sim \text{Gamma}(r, \lambda)$

$$
f_X(x) = \dfrac{\lambda^r}{\Gamma(r)}x^{r-1} e^{-\lambda x}\text{ for } x>0, \lambda>0, \Gamma(r) = (r-1)!
$$

$$
F_X(x) = \left\{
        \begin{array}{ll}
            0 & \quad x<0 \quad \\
            1 - e^{-\lambda x}\displaystyle\sum_{j=0}^{r-1}\dfrac{(\lambda x)^j}{j!}  & \quad x\geq0 \\
        \end{array}
    \right.
$$

$$\text{E}(X) = \dfrac{r}{\lambda}\text{, }\text{ Var}(X) = \dfrac{r}{\lambda^2}$$

Common to see $\alpha = r$ and $\beta = \lambda$

## Identifying gamma RV from word problems

-   Gamma distribution with $r=1$ is same as exponential

    -   Just like Negative Binomial with $r=1$ is same as the geometric distribution

-   Similar to exponential

    -   Look for time between or until events/successes
        -   BUT now we are measuring time until more than 1 success
    -   Look for a rate of the events over time period

## Helpful R code

Let's say we're sitting at the bus stop, measuring the time until 4 buses arrive. We know the bus comes every 10 minutes on average.

-   If we want to know the probability that the 4 buses arrive in the next 50 minutes:

    ```{r}
    #| echo: true
    pgamma(q = 50, rate = 1/10, shape = 4)
    pgamma(q = 50, scale = 10, shape = 4)
    ```

-   If we want to know the time, say $t$, where the probability of the 4 buses arriving at $t$ or earlier is 0.35:

    ```{r}
    #| echo: true
    qgamma(p = 0.35, rate = 1/10, shape = 4)
    ```

-   If we want to know the probability that the 4 buses arrives between 30 and 50 minutes:

    ```{r}
    #| echo: true
    pgamma(q = 50, scale = 10, shape = 4) - pgamma(q = 30, scale = 10, shape = 4)
    ```

-   If we want to sample 20 arrival times for the 4 buses:

    ```{r}
    #| echo: true
    rgamma(n = 20, scale = 10, shape = 4)
    ```

## Remarks

-   The parameter $r$ in a Gamma($r$,$\lambda$) distribution does NOT need to be a positive integer

    -   $r$ is usually a positive integer

-   When $r$ is a positive integer, the distribution is sometimes called an Erlang($r$,$\lambda$) distribution

     

     

-   When $r$ is any positive real number, we have a general gamma distribution that is usually instead parameterized by $\alpha>0$ and $\beta>0$, where:

    -   $\alpha = \text{shape parameter}$ : same as $k$, the total number of events we must witness

        -   In R code example: 4 buses to wait for

    -   $\beta = \text{scale parameter}$ : same as $\lambda$, the rate parameter

        -   In R code example: 1 bus per 10 minutes (1/10)

## Sending money orders

::: columns
::: {.column width="30%"}
::: example
::: ex-title
Example 1
:::

::: ex-cont
On average, someone sends a money order once per 15 minutes. What is the probability someone sends 10 money orders in less than 3 hours?
:::
:::
:::
:::

## Additional Resource

-   Another helpful site with R code: <https://rpubs.com/mpfoley73/459051>
